Red Wine Exploration by Zhang Guanghui
========================================================

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load all of the packages that you end up using
# in your analysis in this code chunk.

# Notice that the parameter "echo" was set to FALSE for this code chunk.
# This prevents the code from displaying in the knitted HTML output.
# You should set echo=FALSE for all code chunks in your file.

library(ggplot2)
library(dplyr)
library(gridExtra)
library(scales)
library(reshape2)
library(GGally)
library(knitr)
```


```{r echo=FALSE,  warning=FALSE, Load_the_Data}
# Load the Data
redwine <- read.csv("wineQualityReds.csv")

```


   
# Univariate Plots Section
### Study the basic information of the dataset
### a) Rows and colums of dataset
```{r echo=FALSE, message=FALSE, warning=FALSE}
# count cols of dataset
nrow(redwine)
ncol(redwine)
```
### b) str of dataset
```{r echo=FALSE, warning=FALSE}
# count rows of dataset
str(redwine)

```
### c) names of variables in dataset
```{r echo=FALSE, warning=FALSE}
# the feature names
names(redwine)

```

### d) head of dataset
```{r echo=FALSE, warning=FALSE}
# get the first 6 rows of dataset
head(redwine)

```

### e) summary of red wine quality
```{r echo=FALSE, warning=FALSE}
# get the summary information
summary(redwine)
```

--- The quality of red wine ranges from 3 to 8. The mean is 5.636 and the median is 6. 

## Draw univariate plots
## f) histogram plots of all features

Plot the histogram plots of all features. Investigate the characteristics of the distribution of each one. 

### f.1) histogram of red wine quality

```{r echo=FALSE}
## knitr configuration
opts_chunk$set(comment = "", error= TRUE, warning = FALSE, message = FALSE,
               tidy = FALSE, cache = F, echo = T,
               fig.width = 5, fig.height = 4)

## R configuration
options(width = 116, scipen = 5)
```

```{r echo=FALSE, Univariate_Plots}


# histogram of red wine quality
ggplot(aes(x=quality), data=redwine) +
  geom_histogram(binwidth=1, color="black") +
  scale_x_discrete(breaks=seq(3, 8, 1), 
                       labels=seq(3, 8, 1) ) +
  coord_cartesian(xlim=c(3,8)) +
  xlab("quality") +
  ggtitle("Histogram of Red Wine Quality")
  
  
```

Six quality score is observed in the dataset. The histogram shows that most of quality is 5 or 6, while quality==3 and quality ==8 are very few.
how many red wine with quality >=5 and <= 6 ?

```{r echo=FALSE, warning=FALSE}
# get the number rows with quality above 5 and below 6
n_mid <- redwine %>%
  filter(quality>=5 & quality<=6) %>%
  summarize(n=n())

```

```{r warning=FALSE}
# get the percentage of rows with quality above 5 and below 6
as.numeric(n_mid)/nrow(redwine)
```

---  82% of redwine is has quality 5 or 6. It wil be very intersting to invesigate why some wine has distinctively high or low quality.


### f.2) histogram of citric.acid
```{r echo=FALSE, warning=FALSE, message=FALSE}
ggplot(aes(x=citric.acid), data=redwine) +
  geom_histogram(color="black", binwidth=0.05) +
  ggtitle("histogram of citric acid") +
  xlab("citric acid (g/dm^3)")

```

This histogram is positively skewed. The count of citric acid starts to descend after 0.50. Change the binwith to see more detailed information. Count the rows of data with citric acid = 0

```{r, echo = FALSE, warning=FALSE, message=FALSE}
ggplot(aes(x=citric.acid), data=redwine) +
  geom_histogram(color="black", binwidth=0.01) +
  ggtitle("histogram of citric acid")+
  xlab("citric acid (g/dm^3)")

# filter out data with the citic acid equals 0
redwine %>%
  filter(citric.acid==0) %>%
  summarize(n=n())
```

--- Changing binwidth shows that the citric.acid is very dispersed. 132 rows of data has citric acid 0. There are three major peaks at 0, 0.25, and 0.50. 

### f.3) histogram of residual.sugar
```{r echo=FALSE, warning=FALSE}

# histogram of residual sugar, limit the x axis limits
ggplot(aes(x=residual.sugar), data=redwine) +
  geom_histogram(color="black", binwidth=0.1) +
  coord_cartesian(xlim=c(1, 4)) +
  ggtitle("histogram of residual sugar") +
  xlab("residual sugar (g/dm^3)")


```
 
This histogram appears quite normally distributed. Get the summary of the feature residual sugar 

```{r echo =FALSE}
summary(redwine$residual.sugar)
``` 

--- The histogram shows that very few wines have residual sugar over 4. The maximum residual sugar content is as high as 15.5

### f.4) histogram of sulphates
```{r echo=FALSE, warning=FALSE}
ggplot(aes(x=sulphates), data=redwine) +
  geom_histogram(color="black", binwidth=0.1) +
  ggtitle("histogram of sulphates")+
  xlab("sulphates (g/dm^3)")

```

The histogram seems to be normally distributed. Change the bin width and x limits to check details.

```{r echo=FALSE, warning=FALSE}
ggplot(aes(x=sulphates), data=redwine) +
  geom_histogram(color="black", binwidth=0.01) +
  scale_x_continuous(breaks=seq(0.25, 1.25, 0.1), limits=c(0.25, 1.25)) +
  ggtitle("histogram of sulphates") +
  xlab("sulphates (g/dm^3)")

```

--- The hisogram of sulphates show that very red wine has sulphates over 0.95. The major is between 0.35 and 0.95

### f.5) histogram of total.sulfur.dioxide

```{r echo=FALSE, warning=FALSE}
ggplot(aes(x=total.sulfur.dioxide), data=redwine) +
  geom_histogram(color="black", binwidth=5) +
  xlim(0, 200) +
  ggtitle("histogram of total sulfur dioxide") +
  xlab("total sulfur dioxide (mg/dm^3)")


```

The distribution is positively skewed. Log-transform the x axis to see the distribution.

```{r echo=FALSE, warning=FALSE}
ggplot(aes(x=log10(total.sulfur.dioxide)), data=redwine) +
  geom_histogram(color="black", binwidth=.05) +
  ggtitle("histogram of total sulfur dioxide (mg/dm^3)") +
  xlab("total sulfur dioxide (mg/dm^3)")

```

--- The Log-transformed plot appears to be normally distributed. The majority is between 1.0 and 2.25.


### f.6) histogram of alcohol
See the distribution of alcohol with different binwidth.

```{r echo=FALSE, warning=FALSE}
ggplot(aes(x=alcohol), data=redwine) +
  geom_histogram(color="black", binwidth=0.5) +
  ggtitle("histogram of alcohol (binwidth=0.5)") +
  xlab("alcohol (%)")


# change the bin width
ggplot(aes(x=alcohol), data=redwine) +
  geom_histogram(color="black", binwidth=0.1) +
  ggtitle("histogram of alcohol (binwidth=0.1)") +
  xlab("alcohol (%)")



```

--- Changing the binwidth reveals that the histogram plot of alcohol is positively skewed. Log transform the x axis  

```{r echo=FALSE}

ggplot(aes(x=log10(alcohol)), data=redwine) +
  geom_histogram(color="black", binwidth=0.01) +
  scale_x_continuous(limits=c(0.9, 1.2), breaks=seq(0.9,1.2, 0.05)) +
  ggtitle("histogram of alcohol (log10)")
```

--- The log-transformed plot shows that the majority is between 0.95 and 1.1, the peak is around 0.975. 

### f.7) histogram of ph
```{r echo=FALSE, warning=FALSE}

# histogram of ph value
ggplot(aes(x=pH), data=redwine) +
  geom_histogram(color="black", binwidth=0.05) +
  scale_x_continuous(limits=c(2.7,4), 
                     breaks=seq(2.7, 4, 0.1)) +
  ggtitle("histogram of pH")
  
```

--- The histogram plot shows that ph looks more like normal distributed with mean around 3.7, and the majority is between 3.0 and 3.6

# Univariate Analysis

### What is the structure of your dataset?
--- There are 1599 observations of data with 12 features, including "fixed.acidity","volatile.acidity","citric.acid","residual.sugar", "chlorides", "free.sulfur.dioxide", "total.sulfur.dioxide", "density", "pH", "sulphates", "alcohol", "quality". The variable "quality" is an ordered factor variable: quality: 0-10 (worst ----> best). Most of redwines are of quality from 5-7. 

### What is/are the main feature(s) of interest in your dataset?
--- It will be impossible for a single variable to determine the quality of red wine, instead the combination and balance of different variables are the main factors. The quality is based on the smell, color, taste of the wine, with which the most determinant variables must be related. So I think the major features influencing includes citric acid, sulphates, alcohol. 


### What other features in the dataset do you think will help support your investigation into your feature(s) of interest?
--- I also think residual sugar, acidity, and chloride will possibly contribute to the quality of wine. I would like to assume residual sugar would probably contribute most at first. But it needs to be confirmed in the following investigation. 

### Did you create any new variables from existing variables in the dataset?
--- No. I did not create a new variables.

### Of the features you investigated, were there any unusual distributions? Did you perform any operations on the data to tidy, adjust, or change the form of the data? If so, why did you do this?
--- 1. The histogram of residual sugar and sulfates indicates that some wine has a very high sugar content. So I use coord_cartesian to limit the x axis range. 
--- 2. The rightly skewed histogram of total.sulfur.dioxide and alchohol are log-transformed. 


# Bivariate Plots Section
- Split the data into two parts, 
- Subset 1000 rows of data,
- plot ggpair figures

```{r echo=FALSE}
## knitr configuration: http://yihui.name/knitr/options#chunk_options
opts_chunk$set(comment = "", error= TRUE, warning = FALSE, message = FALSE,
               tidy = FALSE, cache = F, echo = T,
               fig.width = 8, fig.height = 8)

## R configuration
options(width = 116, scipen = 5)
```

```{r echo=FALSE, Bivariate_Plots}
set.seed(1836)

pair1 <- data.frame(
          alcohol = redwine$alcohol,
          fixed.acidity =redwine$fixed.acidity,
          volatile.acidity = redwine$volatile.acidity,
          citric.acid = redwine$citric.acid,
          residual.sugar = redwine$residual.sugar,
          quality = redwine$quality
)

pair2 <- data.frame(
          chlorides = redwine$chlorides,
          free.sulfur.dioxide = redwine$free.sulfur.dioxide,
          total.sulfur.dioxide = redwine$total.sulfur.dioxide,
          sulphates = redwine$sulphates,
          quality = redwine$quality
)



ggpairs(pair1[sample.int(nrow(pair1), 1000), ])

ggpairs(pair2[sample.int(nrow(pair2), 1000), ])

```

--- In the ggpair figures, all plots of quality vs. other variabls appear to be horizontal strip. Closer investigation at the scatter plots involving quality is needed. 

### Investigate what are the features influencing quality

--- Create ordered factor variable quality.factor

--- Group redwine by quality.factor

--- Get the mean of other features

The code is as follow:

```{r echo=FALSE}
## knitr configuration
opts_chunk$set(comment = "", error= TRUE, warning = FALSE, message = FALSE,
               tidy = FALSE, cache = F, echo = T,
               fig.width = 5, fig.height = 4)

## R configuration
options(width = 116, scipen = 5)
```


```{r}
redwine$quality.factor <- factor(redwine$quality)

redwine_grouped <- redwine %>%
  group_by(quality.factor) %>%
  summarize(mean_alcohol = round(mean(alcohol), 3),
            mean_vol_acid = round(mean(volatile.acidity), 3),
            mean_fix_acid = round(mean(fixed.acidity),3),
            mean_citric_acid = round(mean(citric.acid), 3),
            mean_sugar = round(mean(residual.sugar), 3),
            mean_chlorides = round(mean(chlorides),3),
            mean_free_sulfur = round(mean(free.sulfur.dioxide),3),
            mean_total_sulfur = round(mean(total.sulfur.dioxide), 3), 
            mean_sulphates = round(mean(sulphates), 3),
            mean_ph = round(mean(pH), 3),
            mean_density = round(mean(density), 3),
            n=n())


```

### 1. quality against alcohol

```{r echo=FALSE, warning=FALSE}

# function for computing mean, DS, max and min values
min.mean.sd.max <- function(x) {
  r <- c(min(x), mean(x)-sd(x), 
         mean(x), mean(x)+sd(x), 
         max(x))
  names(r) <-c("ymin", "lower", "middle", "upper", "ymax")
  r
}

# color the point by mean alcohol
ggplot(aes(x= quality.factor, y= alcohol), 
       data = redwine)+
  geom_jitter()+
  scale_color_brewer()+
  scale_x_discrete() +
  stat_summary(fun.data = min.mean.sd.max, geom="boxplot", alpha=0.9)+
  geom_text(data=redwine_grouped, 
            aes(label = mean_alcohol, y = mean_alcohol-0.12, size=3))+
  ggtitle("quality vs. alcohol") +
  xlab("quality") + ylab("alcohol (%)") +
  theme(legend.position="none")

  

```

--- The scatter plot of quality vs alcohol shows that high alcohol red wine has higher quality score. Red wine below the mean quality has alcohol around 10, while red wine above mean quality has obvious higher alcohol. 

### 2. quality against volatile.acidity and fixed.acidity
```{r echo=FALSE, warning=FALSE}

# plot scatter points
ggplot(aes(y= volatile.acidity, x=quality.factor), 
            data = redwine)+
  geom_jitter()+
  geom_point() +
  stat_summary(fun.data = min.mean.sd.max, geom="boxplot", alpha=0.9)+
  geom_text(data=redwine_grouped, 
            aes(label = mean_vol_acid, y = mean_vol_acid-0.05, size=3))+
  scale_x_discrete() +
  ggtitle("Quality vs. volatile acidity") +
  theme(legend.position="none") +
  xlab("quality ") + ylab("volatile acidity (g/dm^3)")


ggplot(aes(y= fixed.acidity, x=quality.factor), 
             data = redwine)+
  geom_jitter()+
  geom_point() +
  stat_summary(fun.data = min.mean.sd.max, geom="boxplot", alpha=0.9)+
  geom_text(data=redwine_grouped, 
            aes(label = mean_fix_acid, y = mean_fix_acid-0.3, size=3))+
  ggtitle("Quality vs. fixed acidity")+
  theme(legend.position="none")+
  xlab("quality ") + ylab("fixed acidity (g/dm^3)")


```

--- The scatter plot of quality vs volatile acidity not only shows all the volatible acidity of each quality, but also shows the mean volatile acidity. High quality red wines appears to have smaller volatile acidity, except the red wines with highest quality 8. However, the relationship between quality and fixed.acidity is more complex. 


### 3. Plot quality against citric acid
```{r echo=FALSE, warning=FALSE}

# add legend guide with customized title
ggplot(aes(y= citric.acid, x=quality.factor), 
       data = redwine)+
  geom_jitter()+
  geom_point() + 
  ylim(-0.1, 1.1) +
  scale_x_discrete() +
  stat_summary(fun.data = min.mean.sd.max, geom="boxplot", alpha=0.9)+
  geom_text(data=redwine_grouped, 
            aes(label = mean_citric_acid, y = mean_citric_acid-0.05, size=3))+
  ggtitle("Quality vs. citric acid")+
  theme(legend.position="none") +
  xlab("quality ") + ylab("citric acid (g/dm^3)")


```

--- The plot shows that high-quality red wine tends to have high citric acid. But citric acid is not a detrminant factor. There is one point with citric acid 1 but with quality only 4. 


### 4. quality against residual.sugar
```{r echo=FALSE, warning=FALSE}


ggplot(aes(y= residual.sugar, x=quality.factor), 
       data = redwine)+
  geom_jitter()+
  geom_point()+
  stat_summary(fun.data = min.mean.sd.max, geom="boxplot", alpha=0.9)+
  geom_text(data=redwine_grouped, 
            aes(label = mean_sugar, y = mean_sugar-0.5, size=3))+
  ggtitle("quality vs. residual sugar")+
  theme(legend.position="none")+
  xlab("quality ") + ylab("residual sugar (g/dm^3)")


```

--- According to the figure, no much differences among mean residual sugar


### 5. quality against chlorides
```{r echo=FALSE, warning=FALSE}


# use jitter to avoid over plotting
# use red color so the mean number can be seen on quality factor 8
ggplot(aes(y= chlorides, x=quality.factor), 
       data = redwine)+
  geom_jitter()+
  geom_point()+
  scale_x_discrete() +
  stat_summary(fun.data = min.mean.sd.max, 
               geom="boxplot", alpha=0.9)+
  geom_text(data=redwine_grouped, 
            aes(label = mean_chlorides, 
                y = mean_chlorides-0.02, 
                size=3, color="red")) +
  ggtitle("quality vs. chlorides")+
  theme(legend.position="none")+
  xlab("quality ") + ylab("chlorides (g/dm^3)")

```

--- The figure shows that not much differences among the mean of chlorides of each quality. The red wine with quality 8 has comparative smaller chlorides. The red wine with the lowest quality has the largest chlrides content. 



### 6. quality against free.sulfur.dioxide, total.sulfur.dioxide, sulphates
```{r echo=FALSE, warning=FALSE}

# use jitter to avoid over plotting
ggplot(aes(y= sulphates, x=quality.factor), 
             data = redwine)+
  geom_jitter()+
  geom_point()+
  scale_x_discrete() +
  stat_summary(fun.data = min.mean.sd.max, geom="boxplot", alpha=0.9)+
  geom_text(data=redwine_grouped, 
            aes(label = mean_sulphates, y = mean_sulphates-0.06, size=3)) +
  ggtitle("quality vs. sulphates")+
  theme(legend.position="none")+
  xlab("quality ") + ylab("sulphates (g/dm^3)")



ggplot(aes(y= free.sulfur.dioxide, x=quality.factor), 
             data = redwine)+
  geom_jitter()+
  geom_point()+
  scale_x_discrete() +
  stat_summary(fun.data = min.mean.sd.max, geom="boxplot", alpha=0.9)+
  geom_text(data=redwine_grouped, 
            aes(label = mean_free_sulfur, y = mean_free_sulfur-3, size=3))+
  ggtitle("quality vs. free sulfur dioxide")+
  theme(legend.position="none")+
  xlab("quality ") + ylab("free sulfur dioxide (mg/dm^3)")
  


ggplot(aes(y= total.sulfur.dioxide, x=quality.factor), 
             data = redwine)+
  geom_jitter()+
  geom_point()+
  scale_x_discrete() +
  stat_summary(fun.data = min.mean.sd.max, geom="boxplot", alpha=0.9)+
  geom_text(data=redwine_grouped, 
            aes(label = mean_total_sulfur, y = mean_total_sulfur-5, size=3))+
  ggtitle("quality vs. total sulfur dioxide")+
  theme(legend.position="none")+
  xlab("quality ") + ylab("total sulfur dioxide (mg/dm^3)")
  



```

--- The scatter plot of quality against sulphates show that high- quality red wine has higher mean sulphates.    

### 7. quality against pH
```{r echo=FALSE, warning=FALSE}

# draw scatter plot, use jitter to avoid over plotting
ggplot(aes(y= pH, x=quality.factor), 
       data = redwine)+
  geom_jitter()+
  geom_point() +
  scale_x_discrete() +
  stat_summary(fun.data = min.mean.sd.max, geom="boxplot", alpha=0.9)+
  geom_text(data=redwine_grouped, 
            aes(label = mean_ph, y = mean_ph -0.04, size=3))+
  theme(legend.position="none")+
  xlab("quality ") + ylab("pH")

```

--- The scatter plot of quality against ph shows the mean of ph for each quality score is quite the same. So I will not regard it importand factor for quality. 

### 8. quality and density
```{r echo=FALSE, warning=FALSE}

# draw scatter plot, use jitter to avoid over plotting
ggplot(aes(y= density, x=quality.factor), 
       data = redwine)+
  geom_jitter()+
  geom_point()+
  scale_x_discrete() +
  stat_summary(fun.data = min.mean.sd.max, geom="boxplot", alpha=0.9)+
  geom_text(data=redwine_grouped, 
            aes(label = mean_density, y = mean_density -0.0008, size=3))+
  theme(legend.position="none")+
  xlab("quality ") + ylab("density (g/cm^3)")

```


--- The density boxplots show that high quality wine tends to have lower volatile acidity, higher citric acid, higher sulphates, and higher alcohol. No obvious relationship is observed between quality and other features.


# Bivariate Analysis

### Talk about some of the relationships you observed in this part of the investigation. How did the feature(s) of interest vary with other features in the dataset?

--- Scatter plots of quality against different features demonstrate that volatile acidity, citric acid,  sulphates, alcohol are strongly related to quality. In generally, lower volatile acidity, higher citric acid, higher sulphates, and higher alcohol would contribute to higher quality. I used to suspect fixed acidity would influence volatile acidity in the same manner. The comparison of plots indicates that it is not true. My little knowledges about red wine tells me that residual sugar may be a sign of quality. As a matter of fact, scatter plot of quality against residual sugar dose not imply direct relationship. It is found out that ph value is strongly related to citric acid, free acidity and volatle acidity. High citric acid and high free acidity would contribute to low ph value, while high volatile acidty could leads to high ph value. 




### Did you observe any interesting relationships between the other features (not the main feature(s) of interest)?
```{r warning=FALSE}
# check the correlation between volatile acidity and fixed acidity
cor.test(redwine$volatile.acidity, redwine$fixed.acidity)

```

--- According to the above investigation, features, like fixed acidity, residual sugar, chlorides, free sulfur dioxide, total sulfur dioxide, ph, are not strongly related to quality. In the scatter plot of quality against sulfur, one interesting discovery is that even though both sulfur dioxide and sulphates has element sulfur in the aspect of chemical composition, the effect on red wine quality is different. Very high quality and very low quality red wine has smaller sulfur dioxide than middle-ranked red wines. In contrast, higher sulphates usually means high-quality wine. 

### What was the strongest relationship you found?
```{r warning=FALSE, echo=FALSE, message=FALSE, results='hide'}

# get the correlation between quality and other strongly-related features
cor.test(redwine$quality, redwine$volatile.acidity)
# -0.3905578 
cor.test(redwine$quality, redwine$citric.acid)
# 0.2263725 
cor.test(redwine$quality, redwine$sulphates)
# 0.2513971
cor.test(redwine$quality, redwine$alcohol)
# 0.4761663 
```

--- The correlation coefficients between quality and volatile acidity, citric acid,  sulphates, alcohol are -0.3906, 0.2264, 0.2514, 0.4762 respectively. The biggest absolute value is found between alcohol and quality. In the bivariate plotting, it is also found that quality is strongly related to alcohol. I need to further study the coupling effects among these four features, and whether they will strengthen or weaken eath others.  



# Multivariate Plots Section
### Density plot of volatile.acidity, alcohol, citric.acid, sulphates group by quality

--- According to previous analysis, four major features influencing quality are volatile acidity, citric acid, sulphates, alcohol. They are numeric variables. Use cut function to create ordered factor in order to investigate categoricall later. The code is as follow:

```{r  Multivariate_Plots}

# use cut to make categorical variable of volatile acidity
redwine$volatile.acidity.bucket <- 
  cut(redwine$volatile.acidity, breaks=c(0.12,0.39, 0.52, 0.53, 0.64, 1.58))

# use cut to make categorical variable of citric acid
redwine$citric.acid.bucket <- 
  cut(redwine$citric.acid, breaks=c(-0.01, 0.25, 0.50, 0.75, 1.0))

# use cut to make categorical variable of sulphates
redwine$sulphates.bucket <-
  cut(redwine$sulphates, breaks=c(0.33, 0.55, 0.62, 0.658, 0.73, 2.0))

# use cut to make categorical variable of alcohol
redwine$alcohol.bucket <- 
  cut(redwine$alcohol, breaks = c(8.3, 9.9, 11.4, 12.9, 14.9))

```


```{r echo=FALSE}
## knitr configuration
opts_chunk$set(comment = "", error= TRUE, warning = FALSE, message = FALSE,
               tidy = FALSE, cache = F, echo = T,
               fig.width = 6, fig.height = 4)

## R configuration
options(width = 116, scipen = 5)
```

```{r echo=FALSE, warning=FALSE}


# plot density graph of the categorical data just created
ggplot(aes(x=quality, color=factor(volatile.acidity.bucket)),             
       data=redwine) +
  geom_density() +
  guides(color = guide_legend( keywidth= 2, 
                               title="volatile acidity (g/dm^3)"))

  

ggplot(aes(x=quality, color=factor(citric.acid.bucket)), 
             data=redwine) +
  geom_density()+
  guides(color = guide_legend( keywidth= 2, 
                               title="citric acid (g/dm^3)"))


ggplot(aes(x=quality, color=factor(sulphates.bucket)), 
             data=redwine) +
  geom_density()+
  guides(color = guide_legend( keywidth= 2, 
                               title="sulphates (g/dm^3)"))

ggplot(aes(x=quality, color=factor(alcohol.bucket)), 
             data=redwine) +
  geom_density()+
  guides(color = guide_legend( keywidth= 2, 
                               title="alcohol (%)"))



```

--- It is observed in the density plots that wine of medium quality 5 more often fall into the range of low citric acid, low sulphates, and low alcohol than wine of quality 4 or 3. And red wine of quality 5 occur quite often in high volatile acidity. 

### Investigate whether other features strengthen alcohol or not

```{r echo=FALSE, warning=FALSE}


# plot quality against alcohol, colored by sulphates.bucket
ggplot(aes(x=quality, y = alcohol, 
               color = sulphates.bucket), 
       data=redwine) +
  geom_jitter(size=2) +
  scale_x_continuous(breaks=c(3,4,5,6,7,8)) +
  stat_smooth(method="lm", formula = y~x, size=1.0) +
  scale_color_brewer() +
  guides(color = guide_legend( keywidth= 2, 
                               title="sulphates (g/dm^3)"))+
  ylab("alcohol (%)") 

# plot quality against alcohol, colored by volatile.acidity.bucket
ggplot(aes(x=quality, y = alcohol, 
               color = volatile.acidity.bucket), data=redwine) +
  geom_jitter(size=2) +
  scale_x_continuous(breaks=c(3 ,4,5,6,7,8)) +
  stat_smooth(method="lm", formula = y~x, size=1.0)+
  scale_color_brewer()+
  guides(color = guide_legend( keywidth= 2, 
                               title="volatile acidity (g/dm^3)"))+
  ylab("alcohol (%)")

# plot quality against alcohol, colored by citric.acid.bucket
ggplot(aes(x=quality, y = alcohol, 
               color = citric.acid.bucket), data=redwine) +
  geom_jitter(size=2) +
  scale_x_continuous(breaks=c(3 ,4,5,6,7,8)) +
  stat_smooth(method="lm", formula = y~x, size=1.0) +
  scale_color_brewer()+
  guides(color = guide_legend( keywidth= 2, 
                               title="citric acid (g/dm^3)"))+
  ylab("alcohol (%)")

  
```

--- The plot shows that higher quality wine has higher alcohol and also higher sulphates. Sulphates.bucket makes it more clear that high alcoohol could contribute to high quality of red wine. 

### Investigate whether other features strengthen volatile.acidity or not

```{r echo=FALSE, warning=FALSE}

# plot volatile acidity against quality, use jitter to avoid over plotting
ggplot(aes(x =quality, y = volatile.acidity, 
           color = alcohol.bucket), data=redwine) +
  geom_jitter(size=2) +
  stat_smooth(method="lm", formula = y~x, size=1) +
  scale_x_continuous(breaks=c(3 ,4,5,6,7,8))+
  scale_color_brewer()+
  guides(color = guide_legend( keywidth= 2, 
                               title="alcohol (%)"))+
  ylab("volatile acidity (g/dm^3)")

ggplot(aes(x =quality, y = volatile.acidity, 
           color = citric.acid.bucket), data=redwine) +
  geom_jitter(size=2) +
  stat_smooth(method="lm", formula = y~x, size=1) +
  scale_x_continuous(breaks=c(3 ,4,5,6,7,8))+
  scale_color_brewer()+
  guides(color = guide_legend( keywidth= 2, 
                               title="citric acid (g/dm^3)"))+
  ylab("volatile acidity (g/dm^3)")

ggplot(aes(x =quality, y = volatile.acidity, 
           color = sulphates.bucket), data=redwine) +
  geom_jitter(size=2) +
  stat_smooth(method="lm", formula = y~x, size=1) +
  scale_x_continuous(breaks=c(3 ,4,5,6,7,8))+
  scale_color_brewer()+
  guides(color = guide_legend( keywidth= 2, 
                               title="sulphates (g/dm^3)")) +
  ylab("volatile acidity (g/dm^3)")

```

--- The plot shows that low-quality red wine tends to have high volatile acidity, but for red wine with citric acid from 0.75 to 1.0, the affect is insignificant. 

### Investigate whether other features strengthen sulphates or not 

```{r echo=FALSE, warning=FALSE}


ggplot(aes(x =quality, y = sulphates, 
                 color = alcohol.bucket), data=redwine) +
  geom_jitter(size=2)+
  stat_smooth(method="lm", formula = y~x, size=1) +
  scale_x_continuous(breaks=c(3 ,4,5,6,7,8)) +
  ylim(quantile(redwine$sulphates, 0.10), quantile(redwine$sulphates, 0.90)) +
  scale_color_brewer()+
  guides(color = guide_legend( keywidth= 2, 
                               title="alcohol (%)"))+
  ylab("sulphates (g/dm^3)")

ggplot(aes(x =quality, y = sulphates, 
                 color = volatile.acidity.bucket), data=redwine) +
  geom_jitter(size=2)+
  stat_smooth(method="lm", formula = y~x, size=1) +
  scale_x_continuous(breaks=c(3 ,4,5,6,7,8))+
  ylim(quantile(redwine$sulphates, 0.10), quantile(redwine$sulphates, 0.90)) +
  scale_color_brewer()+
  guides(color = guide_legend( keywidth= 2, 
                               title="volatile acidity (g/dm^3)"))+
  ylab("sulphates (g/dm^3)")

ggplot(aes(x =quality, y = sulphates, 
                 color = citric.acid.bucket), data=redwine) +
  geom_jitter(size=2)+
  stat_smooth(method="lm", formula = y~x, size=1) +
  scale_x_continuous(breaks=c(3 ,4,5,6,7,8))+
  ylim(quantile(redwine$sulphates, 0.05), quantile(redwine$sulphates, 0.99)) +
  scale_color_brewer()+
  guides(color = guide_legend( keywidth= 2, 
                               title="citric acid (g/dm^3)")) +
  ylab("sulphates (g/dm^3)")

```

--- The plot shows that high quality red wine has comparatively high sulphates. It holds true across different features. When alcohol is between 11.4 and 12.9, the slope of smooth line is smaller, which means that the increase of red wine quality with higher sulphates is smaller that wine in other ranges of alcohol. When citric acid is in the interval of (0.75, 1), the incease of sulphates would cause the quality to drop. Subseting the datasets find that only six wine falls into that category, so it will be regarded as outiliers. 


### Establish a relationship between features

```{r echo=FALSE, warning=FALSE}
lm(redwine$quality~redwine$alcohol + redwine$volatile.acidity + redwine$citric.acid+redwine$sulphates)

```

--- Linear regression analysis skills are used to set up a relationship between quality and other relevent variables. In the regression model, intercept is 2.65, the coefficient for alcohol, volatile acidity, citric acid, sulphates are 0.309, -1.265, -0.079, 0.6955

# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the investigation. Were there features that strengthened each other in terms of looking at your feature(s) of interest?
--- 1. In the vestigation it is found out that high alcohol will contribute to high quality. High sulphate wil generally makes the effect of alcohol more obvious. 

--- 2. Volatile.acidity negatively affect the quality of red wine. High votalie acid would usually make low qulatity. When colord by citric acid, alcohol, sulphate, The effect is quite positive when volatile acidity is low, but become trivial when volatile is high. 

--- 3. sulphate contributes to quality quite positively across different variables. But there are some outliers when alcohol is between 11.4 and 12.9 and citric acid is between 0.75 and 1.


### Were there any interesting or surprising interactions between features?
--- It is quite interesting, even though alcohol, volatile acidity, citric acid, sulphates contribute to quality, they not all work positively. 

### Did you create any models with your dataset? Discuss the strengths and limitations of your model.
--- I established a linear model to predict the quality based on the strong related features. I could give a quality score with the input of alcohol, sulphates, volatile acidity, sulphates. However, the quality of red wine is the combined evaluation of color, flavour, taste, etc. The content of chemicals is indirectly related to these physical characterstic of the red wine.Besides, in the proposed model, the features are numeric variables and predicted quality is also numeric. The model needs to be revised so it can predict quality in more accurate manner. 



# Final Plots and Summary

### Plot One
```{r echo=FALSE}
## knitr configuration
opts_chunk$set(comment = "", error= TRUE, warning = FALSE, message = FALSE,
               tidy = FALSE, cache = F, echo = T,
               fig.width = 5, fig.height = 4)

## R configuration
options(width = 116, scipen = 5)
```

```{r echo=FALSE,  warning=FALSE,Plot_One}
ggplot(aes(x=quality.factor), data=redwine) +
  geom_histogram(binwidth=0.5, color="black") +
  scale_x_discrete(breaks=seq(3, 8, 1), labels=seq(3, 8, 1)) +
  xlab("red wine quality") +
  ggtitle("Histogram of Red Wine Quality")

```

### Description One

--- The histogram shows that the major of red wine has quality 5 or 6. Few red wines have very high or very low quality score. I continue the exploration by investigate which features are strongly related to quality. 

### Plot Two
```{r echo=FALSE,  warning=FALSE,Plot_Two}

p1 <- ggplot(aes(x=factor(quality), y = volatile.acidity), 
       data=redwine) +
  geom_boxplot() +
  xlab("quality") + ylab("volatile acidity (g/dm^3)")

p2 <- ggplot(aes(x=factor(quality), y = citric.acid), 
       data=redwine) +
  geom_boxplot()+
  xlab("quality") + ylab("citric acid (g/dm^3)")

p3 <- ggplot(aes(x=factor(quality), y = sulphates), 
       data=redwine) +
  geom_boxplot()+
  xlab("quality")+ ylab("sulphates (g/dm^3)")

p4 <- ggplot(aes(x=factor(quality), y = alcohol), 
       data=redwine) +
  geom_boxplot()+
  xlab("quality")+ ylab("alcohol (%)")


grid.arrange(p1, p2, p3, p4, ncol=2)

```

### Description Two

--- The plot shows the boxplot of four major features against quality. It has been observed that volative acidity negatively contributes to quality, on the other hand citric acid, sulphates, and alcohol are positively related to quality. The strong correlation cofficient is found between quality and alcohol. The plots indicate that high quality red wine has obvious higher alcohol. 


### Plot Three
```{r echo=FALSE}
## knitr configuration
opts_chunk$set(comment = "", error= TRUE, warning = FALSE, message = FALSE,
               tidy = FALSE, cache = F, echo = T,
               fig.width = 6, fig.height = 5)

## R configuration
options(width = 116, scipen = 5)
```

```{r echo=FALSE, warning=FALSE, Plot_Three}

cuberoot_trans = function() trans_new('cuberoot', 
                                      transform = function(x) x^2,
                                      inverse = function(x) x^2)

ggplot(aes(y =quality.factor, x = alcohol, 
                 color = sulphates.bucket), data=redwine) +
  geom_point(size = 3, position = 'jitter') +
  ylim(3, 8) +
  scale_y_discrete(breaks=seq(3, 8, 1)) +
  xlim(quantile(redwine$alcohol, 0.01), 
       quantile(redwine$alcohol, 0.95)) +
  scale_color_brewer()+
  guides(color = guide_legend( keywidth= 2, 
                               title="sulphates (g/dm^3)"))+
  xlab("alcohol (%)") + ylab("quality")


```

### Description Three

--- A Very strong relationship has been observed between quality and alcohol. The plot also indicates the content of sulphates strengthen the effect of alcohol. Since quality of red wine is a combined evaluation of many different characteristic of the wine, a linear regression model is established to predict quality using four strongly-related features. 


### Reflection
--- This dataset contains 1599 kinds of red wine. Each wine has 12 features. Firstly, I investigated each individual feature by ploting histograms. As I make each plot, I observe the characterstic of each distribution. Afterwards, I explored the quality of red wine across other features and study which features determine the quality. It is discovered that quality is strongly correlated with alcohol, volatile acidity, citric acid, sulphates. A linear regression model is created to predict quality. According to analysis, density and ph value play insignificant role in quality. It is out of my expecation that residual sugar does not quite contribute to high quaity. It is also surprised that quality is negatively related to volatile acidity, but not quite related to fixed acidity. Another interesting finding is that even though sulphates, fixed sulfur dioxide, and total sulfur dioxide has element sulfur, but only sulphates is clearly related to high quality. The limitation of the linear regression model must be noted. Firstly, most of the features of red wine is related to the content of chemicals. The quality of red wine is the combined evaluation of color, flavour, taste, etc. The content of chemicals is indirectly related to these physical characterstic of the red wine. Secondly, the quality of red wine is rated by experts with score from 0 to 10. In the dataset, about eighty percent of the redwines is rated with score 5 and 6, and very few has lower score and higher score. The quality is too clustered. The linear regression model will be more valid if the dataset could involve more red wines with bad and high quality. Thirdly, the quality of red wine is ordered factor with 10 levels. In the proposed model, the features are numeric variables and predicted quality is also numeric. The model needs to be revised so it can predict ordered factor. 
